### 2016-02-03 Manuscript/SW_early_life_telomeres.Rmd
Redraft edits

Lewis G. Spurgin^1,2^,  Kat Bebbington^1^, Eleanor A. Fairfield^1^, Martijn Hammers^2^, Jan Komdeur^3^, Terry Burke^4^, Hannah, L. Dugdale^3,4^, and David S. Richardson^1,5,^.

@@

**Running head:** Telomeres and silver spoons

###Abstract
Understanding the short and long term costs of individual early-life experiences is fundamental to understanding life-history evolution. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, telomere dynamics may help us quantify individual variation in early-life costs, and enhance our understanding of how poor conditions in early life are related to later-life survival. We tested how telomere dynamics are related to spatiotemporal variation in early-life conditions and later-life survival in the Seychelles warbler (*Acrocephalus sechellensis*), across multiple cohorts spanning 14 years. we found that, in accordance with other studies, telomere length and loss are greatest in early life. However, we also show that juvenile telomere lengths varies markedly among cohorts. We found no evidence that early-life social environment (number of helpers, group size) or ecological conditions (territory quality) were related to telomere length, although we found tentative evidence that telomere length varied among summer and winter breeding season. Finally, we found that increased survival later in life was associated with longer telomeres and reduced telomere shortening rates in early life, but these effects could . Our results highlight the inmportance of cohort effects in studies of telomere length, and suggest that

@@

Seychelles warblers are highly territorial and all territories are mapped during the breeding seasons using detailed observational data of foraging and territorial defence behaviour, and surveyed for territory quality [@Richardson2003b]. Territory quality is calculated based on territory size, foliage cover and insect abundance [@Komdeur1992], and territory quality estimates obtained across years are averaged to obtain a single value for each territory [@Hammers2013]. Cousin is subject to intra- and inter-annual variation in rainfall and food availability, and such island-wide temporal variation may override the effects of individual territory quality. As an estimate of seasonal variation in food availability, we calculated an index of the number of insects across the entire island during each breeding season. This  index represents the average number of insects found per unit leaf area over all surveys carried out on the island in a breeding season.

Each time a bird is caught on Cousin body mass and tarsus length are measured (to the nearest 0.1g and 0.1mm, respectively), and age is confirmed on the basis of eye colour [@Komdeur1991a] and previous captures [@Richardson2003b]. A blood sample (*ca* 25 $\mu$l) is taken from each bird captured via brachial venipuncture, and stored at room temperature in 1 ml of absolute ethanol in a 1.5 ml screw-cap microfuge tube.

#####Molecular methods
For each sample, genomic DNA was extracted from a ~2 mm^2^ flake of preserved blood using the DNeasy Blood and Tissue Kit (Qiagen), following the manufacturer's protocol, with the modification of overnight lysis at 37^o^C and a final DNA elution volume of 80 $\mu$l. DNA concentration and purity were quantified using a NanoDrop 8000 Spectrophotometer (ThermoScientific), and DNA integrity was validated by visualization with ethidium bromide after electrophoresis on a 1.2% agarose gel. Sex was determined using the molecular method outlined by Griffiths *et al.* [-@Griffiths1998].

Telomere lengths were measured in `r nrow(juv)+nrow(Loss)` samples, . Of these, `r nrow(juv)` were taken cross-sectionally from birds caught within their first year of life, between `r min(as.numeric(paste(juv$LayYear)))` and `r max(as.numeric(paste(juv$LayYear)))`. For a subset of first-year birds (n = `r nrow(Loss)` individuals) we had longitudinal data, with an additional sample taken within two years of the original catch.

We measured absolute telomere quantity per diploid genome for all samples using a quantitative PCR (qPCR) assay [@Barrett2012]. Post-qPCR processing of samples was carried out as in Barrett *et al.* [-@Barrett2012], with the following amendments. First, averaging of technical repeats was carried out using custom made R scripts (available as supplementary material). Second, a change in batch of SYBR green forced us to raise the annealing temperature of the telomere reaction from 58^oC to 61^oC. This resulted in consistently higher, but repeatable, CQ values for the telomere reaction, so we implemented a simple correction to ensure that telomere lengths were repeatable across annealing temperatures (supplementary material). Finally, we excluded telomere lengths greater than 30kb, which maximised repeatability of our own data [see also @Bauch2012]. Inter-plate repeatability of final telomere lengths was assessed using the R package rptR [@Schielzeth2011].

For the subset of samples with longitudinal data we calculated a rate of telomere loss between the first-year and adult samples by subtracting adult telomere length from early-life telomere length and dividing this difference by the length of time (in days) between sampling events. To account for regression to the mean effects, we applied a correction based on correlations among samples within individuals following Verhulst *et al.* [-@Verhulst2013].

#####Statistical analyses
We performed all analyses using R version 3.0.1 [@RDevelopmentCoreTeam2011]. We used  general linear mixed models along with model averaging to explore how spatial variation in early-life environmental and social conditions influenced telomere length within cohorts. As explanatory variables we included age class (nestling or fledgling), tarsus length, sex, territory quality, season (summer or winter) and the number of helping and non-helping subordinate birds present in the natal territory. As random effects we included birth year and qPCR plate ID. We first created a full model containing all of the above terms, and a top model set was then defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(Loss)`) we repeated the above analyses of telomere dynamics, replacing telomere length with telomere loss as the response variable. In the analysis of telomere loss we also included initial telomere length as an explanatory variable, as longer telomeres have been shown to decrease in length more rapidly, even after correcting for regression to the mean [@Verhulst2013]. Model averaging was carried out using the MuMIn package (version 1.10.5) in R [@Barton2012].

We used Cox regression to test whether survival was related to individual-level telomere length and rate of loss. We ran these models with and without lay year as a frailty term to test whether any effects on telomere length on survival could be explained by cohort effects.

@@

#####Early-life telomere length and age
Inter-plate repeatability of telomere length, based on `r length(table(dd0[duplicated(dd0$BloodID),]$BirdID))` samples measured at least twice, was 0.78 (CI = 0.73-0.82). Mean ($\pm$ standard error) telomere length in first-year birds was longer than in adults of any age (Fig. 1A). We also found substantial age-related variation in telomere length within the first year of life, with longer telomeres in nestlings compared to fledglings (ANOVA; F = `r round2(anova(lm(TLKB~Fledged,data = juv))[1,4])`; P = `r round3(anova(lm(TLKB~Fledged,data = juv))[1,5])`; Fig. 1B).

#####Cohort-level variation in early-life telomere dynamics
Early-life telomere length varied significantly among breeding seasons (one-way ANOVA, F = `r round2(anova(lm(TLKB~factor(FieldPeriodID),data = juv))[1,4])`; P = `r round3(anova(lm(TLKB~factor(FieldPeriodID),data = juv))[1,5])`), with a tendency toward shorter telomeres in more recent years (linear regression of mean telomere length and lay year, R^2^ = `r round2(summary(lm(TLKBmean~LayYear,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TLKBmean~LayYear,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round3(anova(lm(TLKBmean~LayYear,data = juvseason,na.action=na.exclude))[1,5])`; Fig. 2). Variation in mean telomere length over breeding seasons was not related yearly variation in island-wide food availability (linear regression, R^2^ = `r round2(summary(lm(TLKBmean~Insect,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TLKBmean~Insect,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TLKBmean~Insect,data = juvseason,na.action=na.exclude))[1,5])`).

#####Individual-level variation in early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age class, tarsus length and season (summer vs winter) (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lmer(LogTL~1+(1|FieldPeriodID)+(1|PlateID),data=juv,REML = FALSE))-AICc(lmer(LogTL~Fledged+Tarsus+Season+(1|FieldPeriodID)+(1|PlateID),data=juv,REML = FALSE)))`). All three variables importance in the top model set, but only ageclass had confidence intervals that did not overlap zero (Fig. 3A). Telomere length was slightly longer in juveniles born in winter compared to summmer seasons, and this effect appeared most pronounced in nestlings (although sample size of nestlings born in winter seasons is very small) (Fig. 4A). Tarsus length was negatively related to telomere length, although much of this effect could be explained by ageclass, as fledglings have both longer tarsi and shorter telomeres compared to nestlings  (Fig. 4B).

The longitudinal data showed the same patterns as the cross-sectional data. The top model explaining telomere loss again contained age class, tarsus length and season, although in this instance the top model was only a marginally better fit than the null model ($\Delta$AICc = `r round3(AICc(lmer(TROC~(1|LayYear),data=Loss,REML = FALSE))-AICc(lmer(TROC~Fledged+Tarsus+Season+(1|LayYear),data=Loss, REML = FALSE)),lessthan=F)`). Fledglings lost telomeres at a lower rate compared to nestlings, while birds with longer tarsi and birds born in winter seasons had higher rates of telomere shortening compared to those born in the summer (Fig. 3B). However, it is important to note that sample size for this longitudinal analysis was small and all confidence intervals overlapped zero (Fig. 3B).

@@

A survival model including only early-life telomere length revealed a significant and positive effect on survival (estimate = `r round3(summary(coxph(with(juv,Surv(Lifespan+1,Died)~LogTL)))$coefficients[1])`, P = `r round3(summary(coxph(with(juv,Surv(Lifespan+1,Died)~LogTL)))$coefficients[5])`; Fig. 2). However, when we controlled for cohort effects by including lay year as a frailty term in the model, telomere length was no longer a significant predictor of survival (estimate = `r round3(summary(coxph(with(juv,Surv(Lifespan+1,Died)~LogTL+frailty(LayYear))))$coefficients[1,1])`, P = `r round3(summary(coxph(with(juv,Surv(Lifespan+1,Died)~LogTL+frailty(LayYear))))$coefficients[1,6])`). The longitudinal data showed that the amount of telomere shortening experienced in early life had no effect on survival, regardless of whether we included lay year (estimate `r round3(summary(coxph(with(Loss,Surv(Lifespan+1,Died)~TROC+frailty(LayYear))))$coefficients[1,1])`, P = `r round3(summary(coxph(with(Loss,Surv(Lifespan+1,Died)~TROC+frailty(LayYear))))$coefficients[1,6])`), or excluded it (estimate `r round3(summary(coxph(with(Loss,Surv(Lifespan+1,Died)~TROC)))$coefficients[1])`, P = `r round3(summary(coxph(with(Loss,Surv(Lifespan+1,Died)~TROC)))$coefficients[5])`).

###Discussion
Here we show that telomeres are related to both early-life conditions and later-life survival in the Seychelles warbler. We first show that telomere length and rates of shortening in early life are subject to strong cohort effects, and that cohort-level telomere dynamics can be explained by island-wide temporal variation in food availability. We then show that within cohorts, the social environment (i.e. number of helpers) affects telomere length specifically at the nestling stage. Finally, we show that telomere length in early life predicts later-life survival. These findings have important implications for research on life history evolution in wild populations.

The clearest result from our study is that telomere length varies markedly among cohorts. Very few studies have shown that temporal variation in telomere dynamics occurs in natural populations, and the studies that have done so were limited to just two seasons [@Mizutani2013; @Watson2015]. The long-term Seychelles warbler dataset has allowed us, using many more cohorts than has been used in other studies, to show that temporal variation in telomere dynamics does occur at the population level.

We suspect that the novelty in our finding temporal, environmentally-induced variation in telomere dynamics within a population is more due to a lack of available long-term datasets with telomere screening, rather than the Seychelles warbler being unique. Indeed, the environment on Cousin is benign in comparison with many regions outside the tropics, where populations undergo large fluctuations in size [e.g. @Coulson2001]. If our findings are replicated in other systems and population-level variation in early-life telomere dynamics is common in nature, this has a number of ramifications for our understanding of senescence in natural populations. First, our findings suggest that the telomere dynamics of a population at a given point in time represent a snapshot of a temporally varying process. More research is now needed within and across multiple cohorts and populations to better understand the complexity of telomere evolutionary ecology. Second, our findings raise the interesting prospect of using telomeres as indicators of population health - an approach that would be useful to a broad range of conservation and animal health practitioners. However, further research across populations with different environments and genetic histories is required in order to provide further insight into the usefulness of telomeres as biomarkers of population health.

Despite a large sample size and reasonable repeatability, we found little evidence that any of the social or environmental vairables we measured affect juvenile telomere length in the Seychelles warbler. This is surprising a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival. Future research should therefore examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence in natural populations [@Becker2015; @Asghar2014].

We found tentative evidence that, in addition to age class, tarsus length and season had an effect on juvenile telomere length (Figs 3, 4). The tarsus effect most likely reflects the fact that in passerine birds tarsus length is correlated with age during the nestling stage [@Ricklefs1976]. Telomere loss is most rapid early in life due to ongoing cell replication [@Frenck1998], and a negative correlation between telomere length and body size in early life is therefore expected. Indeed, that age class affects both tarsus length and telomere lengths can be clearly seen in our data (Fig. 4B). The difference in telomere length between birds born in summer and winter seasons is more surprising, especially given the directionality of the effect. We find that birds born in winter seasons having longer telomeres (Fig. 4A), while the opp. One possible explanation for this is that parents in good condition are more likely to breed

>>>>

Lewis G. Spurgin^1,2^,  Kat Bebbington^1^, Eleanor A. Fairfield^1^, Martijn Hammers^3^, Jan Komdeur^3^, Terry Burke^4^, Hannah, L. Dugdale^3,4^, and David S. Richardson^1,5,^.

@@

**Running head:** Early-life telomeres and survival

###Abstract
Understanding the short and long term costs of individual early-life experiences is fundamental to understanding life-history evolution. Telomeres, the protective caps at the ends of chromosomes, shorten in response to oxidative stress, and telomere shortening is correlated with reduced survival. Thus, telomere dynamics may help us quantify individual variation in early-life costs, and enhance our understanding of how poor conditions in early life are related to later-life survival. We tested how telomere dynamics are related to spatiotemporal variation in early-life conditions and later-life survival in the Seychelles warbler (*Acrocephalus sechellensis*), across multiple cohorts spanning 14 years. we found that, in accordance with other studies, telomere length and loss are greatest in early life. We then show that juvenile telomere lengths varies markedly among cohorts, with average telomere length varying by . We found no evidence that early-life social environment (number of helpers, group size) or ecological conditions (territory quality) were related to telomere length, although we found tentative evidence that telomere length varied among summer and winter breeding seasons. Finally, we found that increased survival later in life was associated with longer telomeres in early life, but this effect was at the cohort, rather than individual, level. Our results highlight the inmportance of cohort effects in studies of telomere length.

@@

Seychelles warblers are highly territorial and all territories are mapped during the breeding seasons using detailed observational data of foraging and territorial defence behaviour, and surveyed for territory quality [@Richardson2003b]. Territory quality is calculated based on territory size, foliage cover and insect abundance [@Komdeur1992], and territory quality estimates obtained across years are averaged to obtain a single value for each territory [@Hammers2013]. Cousin is subject to intra- and inter-annual variation in rainfall and food availability, and such island-wide temporal variation may override the effects of individual territory quality. As an estimate of seasonal variation in food availability, we calculated an index of the number of insects across the entire island during each breeding season. This index is calculated as the mean number of insects found per unit leaf area over all surveys carried out on the island in a breeding season.

Each time a bird is caught on Cousin body mass and tarsus length are measured (to the nearest 0.1g and 0.1mm, respectively). Using information on eye colour [@Komdeur1991a] and previous captures [@Richardson2003b], we grouped birds into three age categories: one month old (birds still in the nest), 6 months old (fledglings with light grey eyes) and 10 months old (subadults with light brown eyes). A blood sample (*ca* 25 $\mu$l) is taken from each bird captured via brachial venipuncture, and stored at room temperature in 1 ml of absolute ethanol in a 1.5 ml screw-cap microfuge tube.

#####Molecular methods
For each sample, genomic DNA was extracted from a ~2 mm^2^ flake of preserved blood using the DNeasy Blood and Tissue Kit (Qiagen), following the manufacturer's protocol, with the modification of overnight lysis at 37^o^C and a final DNA elution volume of 80 $\mu$l.  Sex was determined using the PCR-based method outlined by Griffiths *et al.* [-@Griffiths1998]. Prior to telomere analysis, DNA concentration and purity were quantified using a NanoDrop 8000 Spectrophotometer (ThermoScientific), and the following thesholds were applied before samples were included for further analysis: i) DNA concentration must be at least 15 ng $\mu$l^-1^ (based on a mean of three measurements), ii) the 260/280 ratio has to be between 1.8 and 2 and, iii) the 260/230 ratio should be higher than 1.8. DNA integrity was further validated by visualization with ethidium bromide after electrophoresis on a 1.2% agarose gel, and all samples with evidence of DNA degradation were re-extracted or excluded.

All DNA extractions that passed the above criteria were diluted to 3.3 ng $\mu$l^-1^ before telomere measurement. We measured absolute telomere quantity per diploid genome for all samples using a quantitative PCR (qPCR) assay, outlined in detail by Barrett *et al.* [-@Barrett2012], with one amendment. A change in batch of SYBR green forced us to raise the annealing temperature of the telomere reaction from 58^o^C to 61^o^C for the majority of the samples (n = `r nrow(subset(dd,Objective %in% c(7:9)))`; see also below). Post-qPCR processing of samples was carried out as in Barrett *et al.* [-@Barrett2012], with the following amendments. First, averaging of technical repeats was carried out using custom-made R scripts (available as supplementary material) instead of GenEx. Second, the change in annealing temperature resulted in consistently higher, but repeatable, CQ values for the telomere reaction, so we implemented a simple correction to ensure that telomere lengths were repeatable across annealing temperatures (supplementary material). Finally, we excluded telomere lengths greater than 30kb, which maximised repeatability of our own data [see also @Bauch2012]. Inter-plate repeatability of final telomere lengths was assessed using the R package rptR [@Schielzeth2011], and for all subsequent analyses we used mean telomere length per sample where we had repeats.

Telomere lengths were measured using a total of `r nrow(juv)+nrow(Loss)` samples, . Of these, `r nrow(juv)` were taken cross-sectionally from birds caught within their first year of life, between `r min(as.numeric(paste(juv$LayYear)))` and `r max(as.numeric(paste(juv$LayYear)))`. For a subset of first-year birds (n = `r nrow(Loss)` individuals) we had longitudinal data, with an additional sample taken as an adult. For these individuals we calculated the absolute amount of telomere loss between the first-year and adult samples by subtracting adult telomere length from early-life telomere length, as well as a rate of telomere loss by dividing this difference by the length of time (in days) between sampling events.

#####Statistical analyses
We performed all analyses using R version 3.0.1 [@RDevelopmentCoreTeam2011]. Telomere length was log transfomred to fit assumptions of normality. We first explored how telomere length varied within the first year of life and over time using linear regression (telomere length vs age in months), and one-way ANOVA (telomere length vs cohort). With the longitudinal data, we tested how telomeres shorten with age by testing how telomere loss and rate of loss were related to the time interval between sampling events, using linear models. If telomeres are lost at a constant rate from early age, we expect the time between sampling events to be positively and linearly relaited to telomere loss no, and unrelated to rate of loss. If, however, telomeres are lost at a greater rate early in life we expect a non-linear or negative relationship between telomere loss and time between sampling events, and a significant (linear or non-linear) decrease in telomere rate of loss with time interval.

We then used a general linear mixed model approach to explore how spatial variation in early-life environmental and social conditions influenced telomere length within cohorts. Model averaging was carried out using the MuMIn package (version 1.10.5) in R [@Barton2012]. We first created a full model containing the following explanatory variables: age (in months), tarsus length, sex, territory quality, season (summer or winter) and the number of helping and non-helping subordinate birds present in the natal territory. We also included interaction terms between age and all the other variables as telomere dynamics are epxected to vary within the first year of life [@Heidinger2012]. As random effects we included cohort, territory ID and qPCR plate ID. Model selection was then performed and a top model set defined, containing all models with AICc $\leq$ 6 compared to the best supported model [@Burnham2011]. We report model-averaged coefficients, confidence intervals and 'relative importance', which reflects the relative weights of each predictor variable across the top model set. For individuals with longitudinal data (*n* = `r nrow(Loss)`) we repeated the above analyses of telomere dynamics, replacing telomere length with telomere loss as the response variable, and excluding the plate ID random effect (as each measurement of telomere loss was based on two or more measurements, and so run on multiple plates).

We used Cox regression, implemented in the survival package in R [@Therneau2011], to test whether survival was related to individual-level telomere length and rate of loss. For the cross-sectional data we used lifespan (in years) as survival time, while for the longitudinal data we used remaining lifespan from the second sampling event. We ran these models with cohort as a frailty term to exclude cohort effects.

@@

#####Early-life telomere length and age
Inter-plate repeatability of telomere length, based on XX samples measured at least twice, was 0.78 (CI = 0.73-0.82). Telomere length dereased with age within the first year of life (R^2^ = `r round2(summary(lm(LogTL~Agemonths,data=juv))$r.squared)`; F = `r round2(anova(lm(LogTL~Agemonths,data=juv))[1,4])`; P = `r round3(anova(lm(LogTL~Agemonths,data=juv))[1,5])`; Fig. 1A). Similarly, telomere shortening based on longitudinal data was significantly greater than zero in nestlings (one sample t-test; P < 0.001), but not in any other age group (all P > 0.5; Fig. 1B).

#####Cohort-level variation in early-life telomere dynamics
Early-life telomere length varied significantly among breeding seasons (one-way ANOVA, F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = juv))[1,4])`; P = `r round3(anova(lm(LogTL~factor(FieldPeriodID),data = juv))[1,5])`). Separating temporal variaiton in telomere length among age classes revealed no significant cohort effects in chicks (F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 1)))[1,4])`; P = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 1)))[1,5])`; Fig. 2A), but significant effects in fledglings (F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 6)))[1,4])`; P = `r round3(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 6)))[1,5])`; Fig. 2B) and subadults (F = `r round2(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 9)))[1,4])`; P = `r round3(anova(lm(LogTL~factor(FieldPeriodID),data = subset(juv,Agemonths == 9)))[1,5])`; Fig. 2C). Variation in median telomere length over breeding seasons in fledglings and subadults was not related temporal variation in  territory quality (linear regression, R^2^ = `r round2(summary(lm(TL~TQ,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TL~TQ,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TL~TQ,data = juvseason,na.action=na.exclude))[1,5])`), island-wide food availability (R^2^ = `r round2(summary(lm(TL~Insect,data=subset(juvseason,Insect<9),na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TL~Insect,data = subset(juvseason,Insect<9),na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TL~Insect,data = subset(juvseason,Insect<9),na.action=na.exclude))[1,5])`) or population density (R^2^ = `r round2(summary(lm(TL~Density,data=juvseason,na.action=na.exclude))$r.squared)`; F = `r round2(anova(lm(TL~Density,data = juvseason,na.action=na.exclude))[1,4])`; P = `r round2(anova(lm(TL~Density,data = juvseason,na.action=na.exclude))[1,5])`).

#####Individual-level variation in early-life telomere dynamics
The top model explaining within-season variation in early-life telomere length contained age class and the number of non-helping subordinates in the natal territory (Table S1). This model was much better supported than the null model ($\Delta$AICc = `r round2(AICc(lmer(LogTL~1+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE))-AICc(lmer(LogTL~Fledged+NonHelper+(1|FieldPeriodID)+(1|PlateID)+(1|BirdID),data=juv,REML = FALSE)))`). The age class effect was very strong and significant, and was present in every model in the top model set (Fig. 3A). The effect of non-helpers, on the other had did not have higher importance across the top model set than the other variables, and confidence intervals overlapped zero (Fig. 3A). Furthermore the model including non-helpers and age class model was not a significantly better fit than a model including age class alone ($\Delta$AICc = 0.06; Table S1).

The top model explaining telomere loss contained age class and tarsus length, and agian this was a significantly better fit than the null model ($\Delta$AICc = `r round2(AICc(lmer(TROC~1+(1|FieldPeriodID),data=Loss,REML = FALSE))-AICc(lmer(TROC~Fledged+Tarsus+(1|FieldPeriodID),data=Loss,REML = FALSE)))`). Fledglings lost telomeres at a lower rate compared to nestlings, while birds with longer tarsi higher rates of telomere shortening (Fig. 3B).

@@

Telomere length in early life did not affect survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~LogTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~LogTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,4])`), and there was no interaction effect between telomere length and juvenile age on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~LogTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glm(SurvivedNext~LogTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,4])`). To separate out cohort-level telomere length affected survival we calculated cohort-level telomere length (i.e. the mean for each cohort) and added this term to a logistic regression. However, we found no effect of telomere length on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~cohortTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~cohortTL+Agemonths,data=juv,family = 'binomial'))$coefficients[2,4])`), and no interaction with juvenile age (estimate = `r round2(summary(glm(SurvivedNext~cohortTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,1])`; P = `r round2(summary(glm(SurvivedNext~cohortTL*Agemonths,data=juv,family = 'binomial'))$coefficients[4,4])`)

Cox regression also showed no effect of individual-level or cohort-level telomere length on survival later in life (individual-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~LogTL+Fledged)))$coefficients[1,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~LogTL)))$coefficients[1,5])`;  cohort-level: estimate = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Fledged+frailty(LayYear))))$coefficients[2,1])`, P = `r round2(summary(coxph(with(juv9,Surv(Lifespan,Died)~cohortTL+Fledged+frailty(LayYear))))$coefficients[2,6])`).

The longitudinal data showed that the amount of telomere shortening experienced in early life had no effect on survival to adulthood (estimate = `r round2(summary(glm(SurvivedNext~TROC+Agemonths,data=subset(Loss,LayYear<20013),family = 'binomial'))$coefficients[2,1])`; P = `r round2(summary(glm(SurvivedNext~TROC+Agemonths,data=subset(Loss,LayYear<2013),family = 'binomial'))$coefficients[2,4])`), or in later life (estimate `r round3(summary(coxph(with(subset(Loss,LayYear<2009),Surv(RemainingLife2+1,Died)~TROC+Agemonths)))$coefficients[1,1])`, P = `r round3(summary(coxph(with(subset(Loss,LayYear<2009),Surv(RemainingLife2+1,Died)~TROC+Agemonths)))$coefficients[1,5])`).

###Discussion
Here we use the long-term study of a closed population of Seychelles warblers to assess the relationships between early-life conditions, telomere length and survival. We find that while telomere length varies

The clearest result from our study is that telomere length varies among cohorts. Very few studies have shown that temporal variation in telomere dynamics occurs in natural populations, and to our knowledge the studies that have done so were limited to just two seasons [@Mizutani2013; @Watson2015]. The long-term Seychelles warbler dataset has allowed us to show that temporal variation in telomere dynamics does occur at the population level over longer time periods.

We suspect that the novelty in our finding temporal, environmentally-induced variation in telomere dynamics within a population is more due to a lack of available long-term datasets with telomere screening, rather than the Seychelles warbler being unique. Indeed, the environment on Cousin is benign in comparison with many regions outside the tropics, where populations undergo large fluctuations in size [e.g. @Coulson2001]. If our findings are replicated in other systems and population-level variation in early-life telomere dynamics is common in nature, this has potential conseaquences for our understanding of telomere dynamics in natural populations. In particular, our findings suggest that the telomere dynamics of a population at a given point in time represent a snapshot of a temporally varying process. More research is now needed within and across multiple cohorts and populations to better understand how how and why population-level telomere dynamics vary over space and time.

In the Seychelles warbler we found only weak evidence that any of the social or environmental vairables we measured affect juvenile telomere length. This is surprising a poor social and ecological environment is known to be detrimental to juvenile Seychelles warblers, both in terms of oxidative stress and later life survival. Possible explanations for this finding include i) temporal variation in telomere length in our data is so strong that we are unable to detect spatial trends, ii) within-cohort variation in telomere length is explained by a variable that we have not included in our analyses, or iii) that effects are generally weak and levels of noise in our telomere measurments and/or ecological data preclude detection of significant effects. Future research should therefore examine, in a quantitative genetic framework, how genetic and environmental components, and their interactions, affect telomere dynamics and senescence in natural populations [@Becker2015; @Asghar2014].

Our data suggest that, in addition to age class, tarsus length and season had weak effects on juvenile telomere length (Figs 3, 4). The tarsus effect most likely reflects the fact that in passerine birds tarsus length is correlated with age during the nestling stage [@Ricklefs1976]. Telomere loss is most rapid early in life due to ongoing cell replication [@Frenck1998], and a negative correlation between telomere length and body size in early life is therefore expected. Indeed, that age class affects both tarsus length and telomere lengths can be clearly seen in our data (Fig. 4B). The difference in telomere length between birds born in summer and winter seasons is more surprising, especially given the directionality of the effect. We find that birds born in winter seasons having longer telomeres (Fig. 4A), while the opp. One possible explanation for this is that parents in good condition are more likely to breed
